### 2015-05-12 background.tex
Edits to code example.

\begin{lstlisting}[caption = Data races in a Linux driver that we discovered using \whoop., label = fig:data_race_example]
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  switch (origin) {
    case 0:
      break;
    case 1:
      offset += file->f_pos; // racy
      break;
    case 2:
      offset += nvram_len;
      break;
    default:
      offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset; // racy
  return file->f_pos; // racy
}
\end{lstlisting}

The most common technique for eliminating data races is using a \emph{locking} mechanism. Locking a set of program statements that access a shared resource creates a \emph{critical section}, i.e. source code that cannot be executed by more than one thread simultaneously. Example~\ref{fig:lock_example} shows how to use a lock to eliminate the data race in Example~\ref{fig:data_race_example}. The example is tricky, because the return statement can potentially race on the \texttt{f\_pos} field. To fix this issue, we store the result in a temporary variable, inside the critical section.

\begin{lstlisting}[caption = Using a lock to eliminate the data races., label = fig:lock_example]
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  mutex_lock(&nvram_mutex); // lock
  switch (origin) {
    case 0:
      break;
    case 1:
      offset += file->f_pos;
      break;
    case 2:
      offset += nvram_len;
      break;
    default:
      offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset;
  loff_t res = file->f_pos; // store result
  mutex_unlock(&nvram_mutex); // unlock
  return res;
}
\end{lstlisting}

@@

>>>>

\begin{figure}[t]
\begin{lstlisting}
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  switch (origin) {
    case 0: break;
    case 1: offset += file->f_pos; break; // racy
    case 2: offset += nvram_len; break;
    default: offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset; // racy
  return file->f_pos; // racy
}
\end{lstlisting}
\caption{Data races in a Linux driver that we discovered using \whoop}\label{fig:data_race_example}
\end{figure}

The most common technique for eliminating data races is using a \emph{locking} mechanism. Locking a set of program statements that access a shared resource creates a \emph{critical section}, i.e. source code that cannot be executed by more than one thread simultaneously. Figure~\ref{fig:lock_example} shows how to use a lock to eliminate the data race in Figure~\ref{fig:data_race_example}. The example is tricky, because the return statement can potentially race on the \texttt{f\_pos} field. To fix this issue, we store the result in a temporary variable, inside the critical section.

@@

\begin{figure}[t]
\begin{lstlisting}
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  mutex_lock(&nvram_mutex); // lock
  switch (origin) {
    case 0: break;
    case 1: offset += file->f_pos; break;
    case 2: offset += nvram_len; break;
    default: offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset;
  loff_t res = file->f_pos; // store result
  mutex_unlock(&nvram_mutex); // unlock
  return res;
}
\end{lstlisting}
\caption{Using a lock to eliminate the data races}\label{fig:lock_example}
\end{figure}
